Conflict Analysis for Mixed-integer SDPs

### 2022-08-02 tex/ConflictAnalysis.tex
Extensions

In order to produce conflict constraints, one needs solutions for the
primal of the LP-relaxation~\eqref{SDP-P}. Because SCIP-SDP preprocesses
certain SDP problems, we briefly discuss how primal solutions can be
obtained in these cases.

@@

>>>>

In order to produce conflict constraints, one needs solutions for the
primal of the LP-relaxation~\eqref{SDP-P}. Because SCIP-SDP preprocesses
certain SDP problems, we briefly discuss how primal solutions can be
obtained in these cases. We start with a general discussion and then
discuss how such solutions can be obtained from the semi-smooth Newton
algorithm for solving SDPs with one variable.

@@

% -------------------------------------------------------------------------
\subsubsection{Obtaining Results from SDP with One Variable}

The semi-smooth Newton algorithm for solving SDPs with one variable can
also produce the corresponding solutions, see~\cite{MatP22}.

\paragraph{Infeasibility I}

The implementation can determine one kind of infeasibility
of~\eqref{eq:basic} by computing a unit eigenvector $w$ for a minimal
eigenvalue $\lambda$ of $A u - B$. If $\lambda < 0$ and $w\T A w > 0$, then
the problem is infeasible.

In this case, we can use $(X^*, \alpha^*, \beta^*) = (w w\T, 0, w\T A w)$
as a ray. To see this, note that
\[
\skal{A}{X^*} + \alpha^* - \beta^* = w\T A w - w\T A w = 0.
\]
Moreover, the objective function satisfies
\[
\skal{B}{X^*} + \ell \alpha^* - u \beta^* = w\T B w - u\, w\T A w = w\T (B -
u A) w = -\lambda\, w\T w > 0,
\]
and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.

\paragraph{Infeasibility II}

A second case of infeasibility of~\eqref{eq:basic} arises if a unit
eigenvector $v$ for a minimal eigenvalue $\lambda$ of $A \ell - B$
satisfies $\lambda < 0$ and $v\T A v \leq 0$.

In this case, we can use $(X^*, \alpha^*, \beta^*) = (v v\T, -v\T A v, 0)$
as a ray. To see this, note that
\[
\skal{A}{X^*} + \alpha^* - \beta^* = v\T A v - v\T A v = 0.
\]
Moreover, the objective function satisfies
\[
\skal{B}{X^*} + \ell \alpha^* - u \beta^* = v\T B v - \ell\, v\T A v = v\T (B -
\ell A) v = -\lambda\, v\T v > 0,
\]
and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.

\paragraph{Infeasibility III}

The third case of infeasibility arises if the iteration terminates with
$\mu \define \mu_k$ and $(v^k)\T A v^k = 0$, where $v \define v^k$ is a
unit eigenvector of $A \mu - B$ corresponding to eigenvalue
$\lambda \define \lambda_k < 0$. In this case, we can take
$(X^*, \alpha^*, \beta^*) = (v v\T, 0, 0)$. To see this, first note that
$(A \mu - B)v = \lambda v$. This implies that
$\mu\, v\T A v - v\T B v = \lambda$, i.e., $- v\T B v = \lambda < 0$. Then
\[
\skal{A}{X^*} + \alpha^* - \beta^* = v\T A v = 0.
\]
Moreover, the objective function satisfies
\[
\skal{B}{X^*} + \ell \alpha^* - u \beta^* = v\T B v > 0,
\]
and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.

\paragraph{Optimality}

Let $\mu$ and $v$ be the final values of the algorithm in the case that no
infeasibility was determined. Then $v$ is a unit eigenvector of $A \mu - B$
corresponding to eigenvalue $\lambda$. Thus, the algorithm terminated
because the condition in the while loop is not satisfied anymore, i.e.,
$\lambda \geq 0$ or $v\T A v \leq 0$. Since the problem is feasible, we
always have $\lambda \geq 0$ and in exact arithmetic the algorithm would
stop with $\lambda = 0$.

First assume that $\delta \define v\T A v > 0$. Note that $\lambda = 0$
implies $(A \mu - B)v = 0$ and thus $v\T B v = \mu\, v\T A v$. Then we can
use $(X^*, \alpha^*, \beta^*) = (\frac{\gamma}{\delta} v v\T, 0, 0)$.  To
see this, note that
\[
\skal{A}{X^*} + \alpha^* - \beta^* = \frac{\gamma}{\delta}\, v\T A v = \gamma.
\]
Moreover, the objective function satisfies
\[
\skal{B}{X^*} + \ell \alpha^* - u \beta^* = \frac{\gamma}{\delta}\, v\T B v
= \frac{\gamma}{\delta}\, \mu\, v\T A v = \gamma\, \mu,
\]
as it should be.

Now assume that $v\T A v = 0$. In this case, strong duality does not need
to hold and it is harder to construct any feasible solution, because we
would need $\skal{A}{X^*} = \gamma$. For instance, if $A$ is negative
semidefinite and $\gamma > 0$, this is impossible to achieve.
